A Cross-Domain Benchmark for Active Learning

### 2024-06-04 doc/al_benchmark_paper.tex
unicode fixes

% TODOs nach Priorität
% Terminologie aufräumen:
% - Query size vs batch size
% - Acq. function vs. AL Alg. vs. AL Method
% - (Un-)encoded vs. semi-supervised/normal
% Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
% Methode zum CD-Diagramm genau erklären (paired-t-test etc.) und die Relevanz fürs paper zeigen
% Advantage-Matrix von Ji et. al. neu aufarbeiten. Wo sind die Unterschiede? Ist das CD-Diagramm besser/schlechter?
% (Un-)Encoded durch Normal uns Semi-Supervised ersetzen, um die Verbindung klar zu machen
% Eventuell Batchsize 1 von den anderen Ergebnissen separieren. Der Vergleich macht mehr Sinn, als einfach die Batchsizes zu mitteln
% Tabelle 1 für Algorithmen statt Domänen. Das zeigt, dass immer entweder BADGE oder Margin (unsere Top 2) fehlt
% Unsere Rankings mit Literatur vergleichen. Wie anders sind wir? -> Evtl. besten Alg. pro Benchmark sammeln und vergleichen
% Causalitäten checken: Viele Dinge werden einfach genannt, ohne das der Grund richtig eingeführt wird
% Fig. 1: Schlechte Performance am Anfang verwirrt nur. Entweder BADGE benutzen, anders plotten, oder im Text in Kontext setzen
% Oracle Forecast im Appendix erklären

@@

Active Learning (AL) deals with identifying the most informative samples for
labeling to reduce data annotation costs for supervised learning tasks. AL
research suffers from the fact that lifts from literature generalize poorly and
that only a small number of repetitions of experiments are conducted. To overcome
this obstacles, we propose \emph{ALBench}, the first active learning benchmark
which includes tasks in computer vision, natural language processing and tabular
learning. Furthermore, by providing an efficient greedy oracle, \emph{ALBench}
can be evaluated with 50 runs for each experiment. We will show, that both the
cross-domain character and the large amount of repetitions are crucial for
sophisticated evaluation of AL research. Concretely, we will show that the
superiority of specific methods varies over the different domains, making it
important to evaluate Active Learning with a cross-domain benchmark.
Additionally, we show, that having a large amount of runs is crucial. With only
conducting five runs as often done in the literature, the superiority of
specific methods can strongly vary with the specific runs. This effect goes so
far, that it may even happen that dedicated methods like \emph{margin-sampling}
are not able to outperform the random baseline.
\end{abstract}

@@

\section{Introduction}
Deep neural networks (NN) have produced state-of-the-art results on many
important supervised learning tasks. Since Deep NNs usually require large
amounts of labeled training data, Active Learning (AL) deals with
selecting the most informative samples out of a large pool of unlabeled data, so
that only these samples need to be labeled. It has been shown that a small
labeled set of this nature can be used to train well-performing models
%\cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}.
% On top of providing a principled way to label unlabeled datasets, active
% learning is one of the two major approaches besides semi-supervised learning to
% make deep learning models more data efficient by requiring only a limited set of
% manually labeled data.
%Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
In the last decade, many different algorithms for AL have been proposed and
almost every method has reported lifts over all its predecessors, \footnote{Out
of all considered algorithms for this paper, only BALD \cite{gal2017deep} did
not claim a new SOTA performance in their result section.} However, real
insights into the current state of AL are hard to draw from these works, due to
the following reasons:
\begin{inparaenum}
\item\label{prob1al} These works do not use a standardized evaluation setting with fixed
datasets and baseline approaches.
\item Due to computational constraints, a lot of works do only a small amount of experimental
runs, hence it is questionable wether the superiority of a specific approach can
be concluded from the conducted experiments.
\item The works are only evaluated in a specific domain, such as computer vision
or language processing. However, AL is a general principle of supervised
learning, and thus methods should be evaluated in multiple domains to assess
their capabilities.
\end{inparaenum}
% %{\let\thefootnote\relax\footnotetext{Code available at: anonymous}}
% %\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
% % (ii) Most of the research is focused on a specific application domain like
% % computer vision, even though AL is a general principle for supervised learning
% % (iii) the reported results can be subject to very high variances across restarts,
% % still A lot of papers do only a hand full of runs, which makes questions the
% % validity of the drawn conclusions.
% % (iv) are not always compared against important baselines like margin
% sampling \cite{wang2014new}.

@@

The applied evaluation protocol in \emph{ALBench} uses 50 runs for
each experiment. By having such a large amount of runs, we can evaluate the
significance of performance gaps and identify the best performing
approaches for each dataset as well as whole domains. 
Furthermore, we show that the small amount of runs other works do in fact 
produce to misleading results. 
To be more specific, we show that if only 3 restarts are employed for each experiment, 
the performance of specific methods strongly varies. As we will see, even the ranking of
the different methods averaged over many datasets fluctuates with the specific set of runs. 
This effect is so strong, that even for a well-established method such
as margin-sampling, it's performance can be significantly better and significantly
worse than random for the same dataset, depending on the seed.

@@

To sum up, \emph{ALBench} is an experimental framework which includes an
efficient oracle approximation, multiple application domains, enough repetitions to
draw valid conclusions and two synthetic tasks to highlight shortcomings of AL methods. 
By being the first benchmark to proving these things in one code-base, 
we believe that \emph{ALBench} is a major step forward of assessing the overall state of Active Learning research, independent of
specific application domains. \emph{ALBench} is publicly available under \emph{anonymous}. \\ [1mm]
Our contributions include the following:
%
\begin{enumerate}
\item\label{cont:repetitions} We show that the small of repetitions that previous works have employed is not
sufficient for meaningful conclusions. Sometimes even making it impossible to assess if a performance is above or below random.
\item\label{cont:oracle} We propose an efficient and performant oracle which is constructed iteratively in a greedy fashion,
overcoming major computational hurdles.
% one of the major obstacles for sufficient repetitions of AL experiments. 
% [Thorben] Das können wir so nicht sagen, glaube ich. Niemand verlangt von uns, dass wir das Oracle auch 50 mal wiederholen
\item\label{cont:domains} We propose \emph{ALBench}, the first general benchmark providing tasks
in the domains of images, text and tabular learning. It further contains
synthetic and pre-encoded data to allow for a sophisticated evaluation of AL
methods. Our experiments show, that there is no clear state-of-the art method
for active learning across different domains.
\item\label{cont:synthdata} We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
designed to hinder active learning by naive decision-boundary- or
clustering-based approaches respectively. Thus, they provide an important tool to identify 
shortcomings of existing AL methods.
\end{enumerate}
%
\begin{table}
	\centering
	\caption{Comparison of our benchmark with the existing literature. Oracle curves
	serve as an approximation of the best possible AL algorithm. Including the
	encoded versions of our datasets we reach 14 datasets, without we have 9. "Semi"
	indicates whether the paper is employing any form of self- or semi-supervised
	learning. A "-" for repetitions means that we could not determine how often each
	experiment is repeated in the respective framework. \emph{ALBench} is the only
	benchmark which reports over enough runs for meaningful results and considers
	all 5 domains.}
	\label{tab:benchmark_comparison}
	\resizebox{\columnwidth}{!}{
\begin{tabular}{lccc|cllll|cl}
Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
\multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
\multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
\multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
\multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
\multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
\multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Few Repetitions are not Sufficient for Meaningful Results}\label{sec:restarts}
To evaluate how many restarts are necessary to obtain conclusive results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
Our best algorithm is margin sampling and we chose the Splice dataset for its average size and complexity. \\
This allows us firstly, to obtain a very strong estimation of the "true" average performance of margin sampling on this dataset and secondly, to draw subsets from this pool of 100 runs.
Setting the size of our draws to $\alpha$ and sampling uniformly, we can approximate a cross-validation process with $\alpha$ restarts.
Each of these draws can be interpreted as a \textbf{reported result in AL literature} where the authors employed $\alpha$ restarts.
%We provide evidence in Fig. \ref{fig:restarts} that previous works might not have evaluated their experiments with a sufficient number of restarts.
Figure \ref{fig:restarts} shows the "true" mean performance of margin sampling (green) in relation to random sampling (black) and the oracle performance (red).
We display 5 random draws of size $\alpha$ in blue.
We can observe that even for a relatively high number of restarts the variance between the samples is extremely high, resulting in some performance curves being worse that random and some being significantly better.
When setting $\alpha = 50$ we observe all samples to converge close to the true mean performance. 
\begin{figure}
        \hspace{-20mm}
	\includegraphics[width=1.24\linewidth]{img/ablation_restarts}
	\caption{Random draws from a pool of 100 runs for margin sampling on the Splice dataset with different numbers of repetitions ($\alpha=\{5,10,50\}$). Green curves are the mean performance of all 100 runs, while the samples are blue. Even with 5 or 10 repetitions, we can observe that single draws for margin sampling display below-random performance (black), while the true mean should be above random.}
	\label{fig:restarts}
\end{figure}
In addition to this motivating example, we carried out our main evaluation (Tab. \ref{tab:results}) multiple times by uniformly sampling 3 random from our available runs and comparing the results.
We found significant differences in the performance of acquisition functions on individual datasets, as well as permutations in the final ranking.
This partly explains the ongoing difficulties in reproducing results for AL experiments and benchmarks.
The details can be found in App. \ref{app:rank_difference}.
For this benchmark we employ 50 restarts of every experiment.
\begin{wraptable}{R}{0.65\textwidth}
	\vspace{-0.7cm}
	\caption{Employed model, chosen budget and available batch sizes for each dataset}
	\vspace{0.1cm}
	\label{tab:batch_sizes}
	{\scriptsize
		\begin{tabular}{l|l|c|c|c|c|c|c|c|c}
			& Model  & B & 1 & 5 & 20 & 50 & 100 & 500 & 1K \\
			\hline
			Enc. DNA    & Linear & 40 & o & o &&&&& \\
			Enc. Splice & Linear & 100 & o & o & o & o &&&\\
			TopV2       & BiLSTM & 200 & o & o & o & o &&& \\
			Splice      & MLP    & 400 & o & o & o & o & o && \\
			DNA         & MLP    & 300 & o & o & o & o & o && \\
			USPS        & MLP    & 400 & o & o & o & o & o && \\
			Enc. Cifar10& Linear & 450 & o & o & o & o & o && \\
			Enc. FMnist & Linear & 500 & o & o & o & o & o && \\
			Enc. USPS   & Linear & 600 & o & o & o & o & o && \\
			News        & BiLSTM & 3K &&& o & o & o & o &\\
			FMnist      & ResNet18& 10K &&&&&& o & o\\
			Cifar10     & ResNet18& 10K &&&&&& o & o \\
		\end{tabular}
	}
	\vspace{-0.55cm}
\end{wraptable}

@@

\bibliographystyle{plain}
\bibliography{al_benchmark_paper.bib}

@@

P3 Query size. Based on our evaluation of the query size we can empirically confirm its importance
with regard to 1) general AL performance and 2) counteracting the cold start problem. The are,
however, surprising findings indicating that the exact interaction between query size and performance
remains an open research question. For instance, we observe the cold start problem for BALD with
Self-SL on CIFAR-100 (∼ 50% accuracy at 2k labeled samples for query sizes of 500 (low-label
regime) and 1k (mid-label regime). On the other hand, in the high label regime (budget of 5k and
query size of 5k) ST and Self-SL models with similar accuracies of 50% and 60%, respectively,
benefit from BALD. Since cold start problems are commonly associated with large query sizes, this
finding seems counter-intuitive, but has been reported before although without further investigations
[6, 20, 43]. To gain a better understanding of this phenomenon and how QMs interact with query
size in the low-label regime, we performed a dedicated experiment series for Self-SL training on
CIFAR-100 and ISIC-2019 (see Figure 3 and Appendix F.2). Due to a significant improvement
of BALD on CIFAR-100 for even smaller query sizes from sub-Random to Random performance
and no worsening on ISIC-2019, we can conclude that small query sizes represent an effective
countermeasure against the cold-start problem for BALD contrary to the findings of [44, 58] and
currently not considered as a solution [6, 20, 43]. This potentially explains the gap between these
findings and more theoretical works which advertise using the smallest query size possible [19, 33].
Importantly, while smaller query sizes appear as a potential solution for the observed instabilities of
BALD, they seem to have no considerable effect on the other QMs. Moreover, our ablation reveals
that even in the low-label settings, BADGE is the overall most reliable of all compared QMs and
exhibits no sub-Random performance. This adds to the existing reports of the robustness of BADGE
for higher label regimes [4, 58].

P4 Classifer configuration. Our results show that method configuration on a properly sized validation
set is essential for realistic evaluation in AL. For instance, our method configuration has the same
effect on the classifier performance as increasing the number of labeled training samples by a factor
of ∼ 5 (e.g. our ST model reach approximately the same accuracy of ∼ 44% trained on 200 samples
compared to models in the AL study by Bengar et al. [6] trained on 1k samples). The effectiveness of
our proposed lightweight HP selection on the starting budget including only three parameters (Sec. 3)
is demonstrated by the fact that all our ST models substantially outperform respective models found
in relevant literature [6, 20, 30, 35, 43, 54, 55] where HP optimization is generally neglected. Details
for this comparison are provided in Appendix H. This raises the question of to which extent reported
AL advantages could have been achieved by simple classifier configurations. Further, our models
also generally outperform expensively configured models by Munjal et al. [44]. Thus, we conclude
that manually constraining the search space renders HP optimization feasible in practice without
decreasing performance and ensures performance gains by Active Learning are not overstated. The
importance of the proposed strategy to optimize HPs on the starting budget for each new dataset is
supported by the fact that the resulting configurations change across datasets.

>>>>

% TODOs nach Priorität
% Terminologie aufräumen:
% - Query size vs batch size
% - Acq. function vs. AL Alg. vs. AL Method
% - (Un-)encoded vs. semi-supervised/normal
% - ThreeClust vs. Honeypot
% Gibt es eine Conclusion für die Encoded-Datasets? Im Moment werden die nicht besprochen
% Methode zum CD-Diagramm genau erklären (paired-t-test etc.) und die Relevanz fürs paper zeigen
% Advantage-Matrix von Ji et. al. neu aufarbeiten. Wo sind die Unterschiede? Ist das CD-Diagramm besser/schlechter?
% (Un-)Encoded durch Normal uns Semi-Supervised ersetzen, um die Verbindung klar zu machen
% Eventuell Batchsize 1 von den anderen Ergebnissen separieren. Der Vergleich macht mehr Sinn, als einfach die Batchsizes zu mitteln
% Tabelle 1 für Algorithmen statt Domänen. Das zeigt, dass immer entweder BADGE oder Margin (unsere Top 2) fehlt
% Unsere Rankings mit Literatur vergleichen. Wie anders sind wir? -> Evtl. besten Alg. pro Benchmark sammeln und vergleichen
% Causalitäten checken: Viele Dinge werden einfach genannt, ohne das der Grund richtig eingeführt wird
% Fig. 1: Schlechte Performance am Anfang verwirrt nur. Entweder BADGE benutzen, anders plotten, oder im Text in Kontext setzen
% Oracle Forecast im Appendix erklären

@@

Active Learning (AL) deals with identifying the most informative samples for
labeling to reduce data annotation costs for supervised learning tasks. AL
research suffers from the fact that lifts from literature generalize poorly and
that only a small number of repetitions of experiments are conducted. To overcome
this obstacles, we propose \emph{ALBench}, the first active learning benchmark
which includes tasks in computer vision, natural language processing and tabular
learning. Furthermore, by providing an efficient greedy oracle, \emph{ALBench}
can be evaluated with 50 runs for each experiment. We show, that both the
cross-domain character and the large amount of repetitions are crucial for
sophisticated evaluation of AL research. Concretely, we show that the
superiority of specific methods varies over the different domains, making it
important to evaluate Active Learning with a cross-domain benchmark.
Additionally, we show, that having a large amount of runs is crucial. With only
conducting five runs as often done in the literature, the superiority of
specific methods can strongly vary with the specific runs. This effect is so strong, that, depending on the seed, even a well-established method's performance can be significantly better and significantly
worse than random for the same dataset. 
\end{abstract}

@@

\section{Introduction}\label{sec:introduction}
Deep neural networks (NN) have produced state-of-the-art results on many
important supervised learning tasks. Since Deep NNs usually require large
amounts of labeled training data, Active Learning (AL) deals with
selecting the most informative samples out of a large pool of unlabeled data, so
that only these samples need to be labeled. It has been shown that a small
labeled set of this nature can be used to train well-performing models
%\cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}.
% On top of providing a principled way to label unlabeled datasets, active
% learning is one of the two major approaches besides semi-supervised learning to
% make deep learning models more data efficient by requiring only a limited set of
% manually labeled data.
%Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
In the last decade, many different algorithms for AL have been proposed and
almost every method has reported lifts over all its predecessors, \footnote{Out
of all considered algorithms for this paper, only BALD \cite{gal2017deep} did
not claim a new SOTA performance in their result section.} However, real
insights into the current state of AL are hard to draw from these works, due to
the following reasons:
\begin{inparaenum}
\item\label{prob1al} These works do not use a standardized evaluation setting with fixed
datasets and baseline approaches.
\item Due to computational constraints, a lot of works do only a small amount of experimental
runs, hence it is questionable wether the superiority of a specific approach can
be concluded from the conducted experiments.
\item The works are only evaluated in a specific domain, such as computer vision
or language processing. However, AL is a general principle of supervised
learning, and thus methods should be evaluated in multiple domains to assess
their capabilities.
\end{inparaenum}
% %{\let\thefootnote\relax\footnotetext{Code available at: anonymous}}
% %\let\thefootnote\relax\footnotetext{Code available at: \url{https://github.com/wernerth94/comparable-active-learning}}
% % (ii) Most of the research is focused on a specific application domain like
% % computer vision, even though AL is a general principle for supervised learning
% % (iii) the reported results can be subject to very high variances across restarts,
% % still A lot of papers do only a hand full of runs, which makes questions the
% % validity of the drawn conclusions.
% % (iv) are not always compared against important baselines like margin
% sampling \cite{wang2014new}.

@@

The applied evaluation protocol in \emph{ALBench} uses 50 runs for
each experiment. By having such a large amount of runs, we can evaluate the
significance of performance gaps and identify the best performing
approaches for each dataset as well as whole domains. 
Furthermore, we show that the small amount of runs other works do in fact 
produce to misleading results. 
To be more specific, we show that if only 3 restarts are employed for each experiment, 
the performance of specific methods strongly varies. As we will see, even the ranking of
the different methods averaged over many datasets fluctuates with the specific set of runs. 
This effect is so strong, that, depending on the seed, even a well-established method's performance can be significantly better and significantly worse than random for the same dataset.

@@

To sum up, \emph{ALBench} is an experimental framework which includes an
efficient oracle approximation, multiple application domains, enough repetitions to
draw valid conclusions and two synthetic tasks to highlight shortcomings of AL methods. 
By being the first benchmark to proving these things in one code-base, 
we believe that \emph{ALBench} is a major step forward of assessing the overall state of Active Learning research, independent of
specific application domains. \emph{ALBench} is publicly available under \emph{anonymous} \todo{Github}. \\ [1mm]
Our contributions include the following:
%
\begin{enumerate}
\item\label{cont:repetitions} We show that the small of repetitions that previous works have employed is not
sufficient for meaningful conclusions. Sometimes even making it impossible to assess if a performance is above or below random.
\item\label{cont:oracle} We propose an efficient and performant oracle which is constructed iteratively in a greedy fashion,
overcoming major computational hurdles.
% one of the major obstacles for sufficient repetitions of AL experiments. 
% [Thorben] Das können wir so nicht sagen, glaube ich. Niemand verlangt von uns, dass wir das Oracle auch 50 mal wiederholen
\item\label{cont:domains} We propose \emph{ALBench}, the first general benchmark providing tasks
in the domains of images, text and tabular learning. It further contains
synthetic and pre-encoded data to allow for a sophisticated evaluation of AL
methods. Our experiments show, that there is no clear state-of-the art method
for active learning across different domains.
\item\label{cont:synthdata} We propose \emph{Honeypot} and \emph{Diverging Sin}, two synthetic datasets
designed to hinder active learning by naive decision-boundary- or
clustering-based approaches respectively. Thus, they provide an important tool to identify 
shortcomings of existing AL methods.
\end{enumerate}
%
\begin{table}
	\centering
	\caption{Comparison of our benchmark with the existing literature. Oracle curves
	serve as an approximation of the best possible AL algorithm. Including the
	encoded versions of our datasets we reach 14 datasets, without we have 9. "Semi"
	indicates whether the paper is employing any form of self- or semi-supervised
	learning. A "-" for repetitions means that we could not determine how often each
	experiment is repeated in the respective framework. \emph{ALBench} is the only
	benchmark which reports over enough runs for meaningful results and considers
	all 5 domains.}
	\label{tab:benchmark_comparison}
	\resizebox{\columnwidth}{!}{
\begin{tabular}{lccc|cllll|cl}
Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
\multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
\multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
\multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
\multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
\multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
\multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Few Repetitions are not Sufficient for Meaningful Results}\label{sec:restarts}
To evaluate how many restarts are necessary to obtain conclusive results in an AL experiment, we computed 100 runs of our top-performing algorithm on one dataset.
Our best algorithm is margin sampling and we chose the Splice dataset for its average size and complexity. \\
This allows us firstly, to obtain a very strong estimation of the "true" average performance of margin sampling on this dataset and secondly, to draw subsets from this pool of 100 runs.
Setting the size of our draws to $\alpha$ and sampling uniformly, we can approximate a cross-validation process with $\alpha$ restarts.
Each of these draws can be interpreted as a \textbf{reported result in AL literature} where the authors employed $\alpha$ restarts.
Figure \ref{fig:restarts} shows the "true" mean performance of margin sampling (green) in relation to random sampling (black) and the oracle performance (red).
We display 5 random draws of size $\alpha$ in blue.
We can observe that even for a relatively high number of restarts the variance between the samples is extremely high, resulting in some performance curves being worse that random and some being significantly better.
When setting $\alpha = 50$ we observe all samples to converge close to the true mean performance. 
\begin{figure}
        \hspace{-20mm}
	\includegraphics[width=1.24\linewidth]{img/ablation_restarts}
	\caption{Random draws from a pool of 100 runs for margin sampling on the Splice dataset with different numbers of repetitions ($\alpha=\{3,5,50\}$). Green curves are the mean performance of all 100 runs, while the samples are blue. Even with 3 or 5 repetitions, we can observe that single draws for margin sampling display below-random performance (black), while the true mean should be above random.}
	\label{fig:restarts}
\end{figure}
In addition to this motivating example, we carried out our main evaluation (Tab. \ref{tab:results}) multiple times by uniformly sampling 3 random from our available runs and comparing the results.
We found significant differences in the performance of acquisition functions on individual datasets, as well as permutations in the final ranking.
This partly explains the ongoing difficulties in reproducing results for AL experiments and benchmarks.
The details can be found in App. \ref{app:rank_difference}.
For this benchmark we employ 50 restarts of every experiment.
\begin{wraptable}{R}{0.65\textwidth}
	\vspace{-0.7cm}
	\caption{Employed model, chosen budget and available batch sizes for each dataset}
	\vspace{0.1cm}
	\label{tab:batch_sizes}
	{\scriptsize
		\begin{tabular}{l|l|c|c|c|c|c|c|c|c}
			& Model  & B & 1 & 5 & 20 & 50 & 100 & 500 & 1K \\
			\hline
			Enc. DNA    & Linear & 40 & o & o &&&&& \\
			Enc. Splice & Linear & 100 & o & o & o & o &&&\\
			TopV2       & BiLSTM & 200 & o & o & o & o &&& \\
			Splice      & MLP    & 400 & o & o & o & o & o && \\
			DNA         & MLP    & 300 & o & o & o & o & o && \\
			USPS        & MLP    & 400 & o & o & o & o & o && \\
			Enc. Cifar10& Linear & 450 & o & o & o & o & o && \\
			Enc. FMnist & Linear & 500 & o & o & o & o & o && \\
			Enc. USPS   & Linear & 600 & o & o & o & o & o && \\
			News        & BiLSTM & 3K &&& o & o & o & o &\\
			FMnist      & ResNet18& 10K &&&&&& o & o\\
			Cifar10     & ResNet18& 10K &&&&&& o & o \\
		\end{tabular}
	}
	\vspace{-0.55cm}
\end{wraptable}

@@

\bibliographystyle{plain}
\bibliography{al_benchmark_paper.bib} 
\begin{itemize}
	\item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
	\item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
	\item Did you include the license to the code and datasets? \answerNA{}
\end{itemize}

\begin{enumerate}

	\item For all authors...
	\begin{enumerate}
		\item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
		\answerYes{See list of contributions in the Sec. \ref{sec:introduction}.}
		\item Did you describe the limitations of your work?
		\answerTODO{}
		\item Did you discuss any potential negative societal impacts of your work?
		\answerTODO{}
		\item Have you read the ethics review guidelines and ensured that your paper conforms to them?
		\answerTODO{}
	\end{enumerate}

	\item If you are including theoretical results...
	\begin{enumerate}
		\item Did you state the full set of assumptions of all theoretical results?
		\answerTODO{}
		\item Did you include complete proofs of all theoretical results?
		\answerTODO{}
	\end{enumerate}

	\item If you ran experiments (e.g. for benchmarks)...
	\begin{enumerate}
		\item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
		\answerTODO{}
		\item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
		\answerTODO{}
		\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
		\answerTODO{}
		\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
		\answerTODO{}
	\end{enumerate}

	\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
	\begin{enumerate}
		\item If your work uses existing assets, did you cite the creators?
		\answerTODO{}
		\item Did you mention the license of the assets?
		\answerTODO{}
		\item Did you include any new assets either in the supplemental material or as a URL?
		\answerTODO{}
		\item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
		\answerTODO{}
		\item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
		\answerTODO{}
	\end{enumerate}

	\item If you used crowdsourcing or conducted research with human subjects...
	\begin{enumerate}
		\item Did you include the full text of instructions given to participants and screenshots, if applicable?
		\answerTODO{}
		\item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
		\answerTODO{}
		\item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
		\answerTODO{}
	\end{enumerate}

\end{enumerate}

@@

P3 Query size. Based on our evaluation of the query size we can empirically confirm its importance
with regard to 1) general AL performance and 2) counteracting the cold start problem. The are,
however, surprising findings indicating that the exact interaction between query size and performance
remains an open research question. For instance, we observe the cold start problem for BALD with
Self-SL on CIFAR-100 (~50\% accuracy at 2k labeled samples for query sizes of 500 (low-label
regime) and 1k (mid-label regime). On the other hand, in the high label regime (budget of 5k and
query size of 5k) ST and Self-SL models with similar accuracies of 50\% and 60\%, respectively,
benefit from BALD. Since cold start problems are commonly associated with large query sizes, this
finding seems counter-intuitive, but has been reported before although without further investigations
[6, 20, 43]. To gain a better understanding of this phenomenon and how QMs interact with query
size in the low-label regime, we performed a dedicated experiment series for Self-SL training on
CIFAR-100 and ISIC-2019 (see Figure 3 and Appendix F.2). Due to a significant improvement
of BALD on CIFAR-100 for even smaller query sizes from sub-Random to Random performance
and no worsening on ISIC-2019, we can conclude that small query sizes represent an effective
countermeasure against the cold-start problem for BALD contrary to the findings of [44, 58] and
currently not considered as a solution [6, 20, 43]. This potentially explains the gap between these
findings and more theoretical works which advertise using the smallest query size possible [19, 33].
Importantly, while smaller query sizes appear as a potential solution for the observed instabilities of
BALD, they seem to have no considerable effect on the other QMs. Moreover, our ablation reveals
that even in the low-label settings, BADGE is the overall most reliable of all compared QMs and
exhibits no sub-Random performance. This adds to the existing reports of the robustness of BADGE
for higher label regimes [4, 58].

P4 Classifer configuration. Our results show that method configuration on a properly sized validation
set is essential for realistic evaluation in AL. For instance, our method configuration has the same
effect on the classifier performance as increasing the number of labeled training samples by a factor
of ~5 (e.g. our ST model reach approximately the same accuracy of ~44\% trained on 200 samples
compared to models in the AL study by Bengar et al. [6] trained on 1k samples). The effectiveness of
our proposed lightweight HP selection on the starting budget including only three parameters (Sec. 3)
is demonstrated by the fact that all our ST models substantially outperform respective models found
in relevant literature [6, 20, 30, 35, 43, 54, 55] where HP optimization is generally neglected. Details
for this comparison are provided in Appendix H. This raises the question of to which extent reported
AL advantages could have been achieved by simple classifier configurations. Further, our models
also generally outperform expensively configured models by Munjal et al. [44]. Thus, we conclude
that manually constraining the search space renders HP optimization feasible in practice without
decreasing performance and ensures performance gains by Active Learning are not overstated. The
importance of the proposed strategy to optimize HPs on the starting budget for each new dataset is
supported by the fact that the resulting configurations change across datasets.
